Visually pleasing knot projections

### 2023-02-02 inst/jma_rebuttal_2023_01_30.tex

The content that is most fitting for a JMA paper is a discussion of
what makes a knot diagram "beautiful"  and that discussion will
clearly be biased by the personal view of the author (perhaps enhanced
by a "user-study" among family and colleagues).  Thus lines 14 - 20 on
page 4 are of prime importance and could be fleshed out more.

The main contribution to JMA would then be a detailed discussion of
how the author has struggled to capture these goals in the form of
mathematical functions that can be optimized and will then approximate
these goals.  In this respect, the manuscript still has some
shortcomings.

They may be fixed in two ways: Enhance the detailed explanations of
the choices made in creating package [27]; and add a more extended
discussion of a few things that might be approached differently in a
future package ( -- which can also handle links).

In view of this, I recommend that Appendices A and B should be firmly
integrated into the main text.  The discussions therein relate rather
directly to what I see as the main contribution of a JMA paper on this
topic.\textcolor{blue}{Done.}

@@

For each crossing point, the user would see an orthogonal "cross-hair"
that can be shifted and rotated as a whole.  In addition, there are
two parameters (handle-pairs) that control the "velocities" of the two
strands through the crossing.  The optimization routine would
fine-tune these velocities to achieve matching curvature for the two
sequentially connected Bezier curves.

And, yes, if a crossing point falls on a line of mirror symmetry, the
crosshair would automatically be angled at 45 degrees with respect to
this line.  And if the knot has n-fold rotational symmetry, n such
cross hairs would be placed at the vertices of a regular n-gon.  (All
this is not such a "nightmare" if it is properly considered from the
beginning.)

I don't expect that anything in the current system should be changed;
but this approach should definitely be considered in a future version
of such a package.
\textcolor{blue}{\\ This scheme certainly sounds achievable and I have
  alluded to it in the manuscript at the two trefoil knot
  discussion.\\ \\}

RE: Mismatched diagram and symmetry specifications: ``This situation
cannot occur" -- If the user makes a mistake, will the current system
send a suitable error message?\textcolor{blue}{\\For convenience, I
  reproduce the original exchange below:\\ \\} {\tiny Reviewer's
  comment, round 3: ``If all nodes fulfill the specified symmetry
  constraints (as shown on page 11, lines 15-22 and on lines 35-43),
  but the connection diagram does not completely adhere to the
  specified symmetry - What happens? - Does the package check this and
  let the user know about the discrepancy?"}  \textcolor{blue}{\tiny
  My original reply: ``This situation cannot occur: symmetric nodes
  imply a symmetric diagram".\\ \\} \textcolor{blue}{A more nuanced
  reply would be to observe that in the event of a mismatch between
  diagram and symmetry specifications, the symmetry specifications
  take priority and the resulting symmetrised diagram automatically
  accommodates the imposed symmetry.  In practice what happens is
  immediately visually apparent: one sees beautiful symmetric nodes
  joined by Bezier curves that do not possess the intended topology.
  The diagram is a disorganised and ugly tangled mess, clearly not
  reflecting the intended configuration.  In practice, one glance at
  the diagram simultaneously reveals the problem and its solution
  (just reconnect the dots in the correct order).\\ \\}

@@

RE: A hierarchical way of dealing with symmetry: ``This is exactly
what I have done!  The referee gives an accurate summary of the code's
methodology." - Place a corresponding, extended description into the
main text.\textcolor{blue}{\\Done.  New text added in section 3.2:
  From a numerical perspective, the software takes advantage of a
  hierarchical representation of symmetrical knots: the curve is
  specified for one unique fragment of the knot, and the complete knot
  is reconstructed by replicating that fragment with mirroring and
  rotation as required.  The optimization is effectively carried out
  for this unique fragment (with proper end-conditions).\\ \\}

RE: A slider interface for changing some optimization parameters: It
is clear why this would not be practical in the current system, where
the optimization of a pretzel knot ``may take a few weeks."  But this
also suggests that there is much room for improvement!  Perhaps some
badness functions are not really needed or are badly implemented.
\textcolor{blue}{\\ This is an excellent suggestion and indeed many of
  the badness functions may well turn out to be superfluous, or at
  least susceptible to faster heuristics [although I would argue that
    the functions all adhere to efficient vectorised R coding
    standards].  If it turns out that some of the badness weightings
  can be set to zero without degrading the resulting visual
  attractiveness, this would certainly be worth pursuing.\\ \\}

RE: Link projections: In a future system, the planning for handling
links should come in right at the beginning.  It is not at all clear
why "Each component would have its own badness."

Many of the important badness functions (e.g. separations of nodes or
of non-crossing curve segments) are based purely on local geometry and
need not know what component of a link they refer to.  Perhaps the
overall length of a knot or of a link-component is not a good metric,
and should be replaced with some other functions of the length of
curve segments between crossings.  \textcolor{blue}{I have spent many
  hours pondering this comment, in particular my original assertion
  that each component would have its own badness.  I took this to be
  axiomatic, but on [extensive] reflection the referee makes an
  excellent point, specifically that the badness metrics are all {\em
    local}.  This is a profound observation, one that eluded me over
  several years of developing and using the software and thinking
  about link diagrams.  I think I missed it because the badness
  functions are largely geared towards preserving the {\em global}
  topology of the diagram [including preventing Reidemeister moves]
  and it seemed natural to consider a link diagram from a topological
  perspective, that is, to consider the badness of each link
  separately.  But the referee's comment is infuriatingly obvious once
  it has been pointed out.  Much of the appendix B is irrelevant in
  the light of this observation; I have shortened it considerably and
  rephrased to reflect the new thinking.  One suggestion above was to
  incorporate appendices A and B in to the main body of the text,
  which I have done, but to accommodate this comment I have
  restructured the text (of what was appendix B) to reflect this
  insight.\\ \\}

### 2023-02-02 inst/jma_knot.tex
fixes #87

Further work might include functionality to deal with links.  Much of
the required functionality is (in principle) already present in the
software.  Each component would have its own badness, and in addition
there would be $n(n-1)/2$ inter-component interaction terms measuring
features such as closeness between non-intersecting strands of
different components.  However, the implementation of such
functionality is not straightforward and several problems stand out.
Firstly, dealing with inter-component intersection points is
difficult.  Currently, crossing points are represented as a
(symmetrical) Boolean matrix with rows and columns corresponding to
strands, and entry $(i,j)$ being whether strand $i$ intersects with
strand $j$.  The link generalization would be a symmetric matrix with
entries that are themselves intersection matrices: an object with four
indices and entry $(i,j,k,l)$ indicating whether strand $j$ of
component $i$ intersects strand $l$ of component $k$.  Such objects
are unwieldy to work with and are typically difficult to manipulate.
Secondly, the imposition of symmetry is not straightforward in the
case of links: one might desire that certain components of a link have
a particular kind of symmetry and others to have a different kind (or
no!) symmetry.  Further, one unexpectedly difficult problem was
dealing with a subset of components that individually possess no
symmetry but collectively possess mirror (or rotational) symmetry.
Even the simplest links, as in Figure~\ref{thistlethwaite}, can be
problematic.  Consider Thistlethwaite's L4a1, for example: one would
expect the two components to be identical except for a 90-degree
rotation, and implementing this in the context of the package does not
seem to be at all easy.  In general, new badnesses would be needed.
Further, in L7a1, the smaller component should be not only as circular
as possible (?), but also one might desire the overall length to be
smaller than that of the other component.  It would be reasonable to
conclude that dealing with links might be possible in principle but,
due to a number of technical and mathematical reasons, considerably
harder than the single-component links considered here.

@@

\clearpage
\section{Gallery}

>>>>

Further work might include functionality to deal with links.  Much of
the required functionality is (in principle) already present in the
software.  Although it might be tempting to assign each link component
its own badness, and to combine these in some way along with some form
of inter-component interaction term, a perceptive referee observed
that the badness components are all based purely on {\em local}
geometry.  As such, many of the software and mathematical principles
designed for knots will apply similarly to links with little
modification.  Consider Thistlethwaite's L4a1
(Figure~\ref{thistlethwaite}), for example: one would expect the two
components to be identical except for a 90-degree rotation, and
implementing this in the context of the package is straightforward in
principle: the symmetry object does not depend on the overall linking
topology of the diagram.  However, dealing with a subset of components
that individually or collectively possess mirror (or rotational)
symmetry, can be problematic.  Consider L7a1, in which only the
smaller component should possess circular symmetry; such restrictions
would require substantial work to implement.

@@

\section{Gallery}
